The Diminishing Returns of Music in the Digital Era

### 2024-04-17 paper/paper.qmd
data and model sections done

library(tidyverse)
library(knitr)
library(readr)
library(ggplot2)
library(palmerpenguins)
```

@@

## Source {#sec-source}
The main data source is a dataset from the Recording Industry Association of America (RIAA) (@RIAA), tracking recorded music revenues by format in the United States from 1973 to 2019. This dataset was generated by Matt Bass and has been adjusted for inflation, offering a standardized monetary comparison across five decades of music industry economics. Since the RIAA have yet to properly respond to requests for the use of their dataset, I made use of a Kaggle dataset from Kaggle user "Larxel" that cites the RIAA dataset.

@@

The RIAA dataset published by @RIAA did not require any further data cleaning. @tbl-averagerevenue shows how in this sample, the most valuable format of music consumption is the CD. This is because CDs are cheap to produce and there is a larger return on investment because they are physical format which have a larger cost than digital formats to the consumer. The least valuable format of music consumption is the music video. This may be in part due to the fact that music videos are immensely costly and they are slowly losing traction due to the rise in social media.
```{r}
#| echo: false
#| eval: true
#| label: tbl-averagerevenue
#| tbl-cap: "Music Industry Average Revenue per Format"
#| warning: false
#| results: asis
#| 
music_data <- read_csv("/Users/victortimir/music-industry-analysis/data/raw_data/musicdata-1.csv")

print(average_revenue_table)
```
\newpage

```{r}
#| echo: false
#| eval: true
#| label: tbl-revenuechange
#| warning: false
#| results: asis
#| 
music_data <- read_csv("/Users/victortimir/music-industry-analysis/data/raw_data/musicdata-1.csv")

@@

@@

Talk way more about it.

@@

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

>>>>

library(tidyverse)
library(knitr)
library(readr)
library(ggplot2)
library(GGally)
library(palmerpenguins)
```

@@

## Source {#sec-source}
The main data source is a dataset from the Recording Industry Association of America (RIAA) (@RIAA), tracking recorded music revenues by format in the United States from 1973 to 2019. This dataset was generated by Matt Bass and has been adjusted for inflation, offering a standardized monetary comparison across five decades of music industry economics. Since the RIAA have yet to properly respond to requests for the use of their dataset, we made use of a Kaggle dataset from Kaggle user "Larxel" that cites the RIAA dataset.

@@

The RIAA dataset published by @RIAA was cleaned to account for the irrelevence of the "number_of_records" column. Using the dataset, we were able to replicate all of the means and standard deviations of the music formats, as shown in @tbl-summary1. The summary statistics illustrate how the high standard deviations across the board reflect a wide variation in yearly revenue, which could be attributed to a certain format's lifecycle or other market changes over time. For example, 8-Track tapes, despite their irrelevance, have an intriguingly high mean revenue.

However, newer formats like digital downloads present lower mean revenues and relatively lower standard deviations, indicating a more consistent performance year-over-year but at a lower profitability scale. This could reflect market saturation (which is growing exponentially in the age of the internet), pricing strategies, or the competition from streaming services. The disparity between these formats highlights the evolution of music consumption, with digital formats overtaking physical mediums and reflecting changing consumer preferences and the influence of technological advancements such as social media and streaming services on the music industry's revenue streams.

```{r}
#| label: tbl-summary1
#| tbl-cap: Summary Statistics of the Music Formats
#| echo: false
#| warning: false
#| message: false
#| results: asis

music_data <- read_csv("/Users/victortimir/music-industry-analysis/data/raw_data/musicdata-1.csv")

summary_stats <- music_data %>%
  group_by(format) %>%
  summarise(
    mean_revenue = mean(value_actual, na.rm = TRUE),
    sd_revenue = sd(value_actual, na.rm = TRUE)
  ) %>%
  ungroup()

kable(summary_stats, caption = "Summary Statistics for Music Formats")
```

@fig-revenuechange illustrates the year-over-year percentage change in revenue for various music formats, demonstrating the music industry's dynamic nature over time. For example, the sharp volatility observed in the cassette format's line is indicative of its rise and fall in popularity, reflective of technological obsolescence and subsequent market shifts. CDs exhibit less volatility but show a general decline in recent years, which correlates with the proliferation of digital music. The download single format and LP/EP show resurgence possibly due to niche market appeal and the revival of vinyl records among collectors and other enthusiasts. These revivals can be traced to social media trends as well, further propagating the internet's immense influence on the music industry. The paid subscription format presents a growing trend, emphasizing the industry's shift towards streaming services as the primary source of music consumption. This graph not only highlights broader trends in the industry, but it also lines up the death of physical media with the rise of the internet, which have reshaped the way music is distributed and monetized. The main concern of the contemporary music industry is also present in this graph as the most consumed format of music (paid subscriptions) are declining in revenue year by year.

```{r}
#| echo: false
#| eval: true
#| label: fig-revenuechange
#| fig-cap: Revenue Change for the Top Music Formats
#| warning: false
#| results: asis

music_data <- read_csv("/Users/victortimir/music-industry-analysis/data/raw_data/musicdata-1.csv")

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-averagerevenue
#| tbl-cap: "Music Industry Average Revenue per Format"
#| warning: false
#| results: asis
#| 
music_data <- read_csv("/Users/victortimir/music-industry-analysis/data/raw_data/musicdata-1.csv")

print(average_revenue_table)
```

@tbl-averagerevenue shows how in this sample, the most valuable format of music consumption in North America is the CD. This is because CDs are cheap to produce and there is a larger return on investment because they are physical format which have a larger cost than digital formats to the consumer. The least valuable format of music consumption in North America is the music video. This may be in part due to the fact that music videos are immensely costly and they have been slowly losing traction due to the rise in social media and streaming services.

@@

@@

The goal of our modelling strategy is twofold. Firstly, we seek to evaluate how different music formats contribute to overall industry revenue over time. This involves analyzing the changes in revenue streams from physical sales such as CDs and vinyl to digital formats like streaming and downloads. Secondly, we seek to use these insights to forecast future revenue trends for these formats.

In the section below, we outline the Bayesian analysis model utilized to examine the linear regression model of revenue generation per music format.

@@

Define $y_i$ as the annual revenue from a given music format. Then $\beta_i$ is the effect of physical sales, $\gamma_i$  is the effect of digital downloads, $\theta_i$ is the effect of streaming subscriptions, and $\phi_i$ is the effect of ad-supported streaming. Let $x_i$ be a vector containing the presence of each format in a given year.This is all measured in USD.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i x_{i1} + \gamma_i x_{i2} + \theta_i x_{i3}\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\theta &\sim \mbox{Normal}(0, 2.5) \\
\phi &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

@@

We expect a positive relationship between physical formats like CDs and vinyl records and revenue in the earlier years, reflecting the technology and consumer habits of those times. As we move into the digital age, we anticipate seeing a shift with digital downloads and streaming services gaining prominence and thus a positive relationship with revenue in the later years of the dataset.

Particularly, we predict that the rise of streaming services will correlate with a decrease in revenue from physical formats due to the convenience and accessibility of music online. Additionally, the advent of certain ad-supported streaming platforms might introduce a more complex relationship with revenue, potentially showing growth in consumer numbers but a slower rate of revenue increase due to its reliance on ad sales rather than direct consumer purchases.

Interactions between different formats are also of interest. For example, the coexistence of digital downloads and streaming services may reveal competitive or complementary dynamics. A competitive interaction might exhibit a negative relationship between the revenue of one format as the other rises, while a complementary interaction may not show a significant negative impact. These dynamics are crucial for understanding the evolution of music consumption and the financial viability of different formats in the music industry.
